Supp figure 4

### 2021-12-02 Figures/Supp_fig4.Rmd
Refactor Supp_fig4.Rmd

```{r, echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(reshape2)
  library(ggpubr)
  library(tidyverse)
  library(Hmisc)
  library(ggridges)
  library(gridExtra)
  library(grid)
})
knitr::opts_chunk$set(fig.width=15, fig.height=15) 
```

@@

tcga_deconv = read.table("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/deconvolutions_pipeline/all_deconvolutions/all_deconvolutions_all_TCGA.txt", header=T)
Aran_all_facs <- read.csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Aran_facs.csv", sep = "\t")
HE_all_TCGA <- read.csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv", row.names = 1)

colnames(tcga_deconv) = colnames(tcga_deconv) %>% str_replace_all("(CBSX__Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix_)|(CBSX_Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix)", "CBSX.NSCLC")
colnames(tcga_deconv) = colnames(tcga_deconv) %>% str_replace_all("(CBSX__scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b_)|(CBSX_scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b)", "CBSX.melanoma")
colnames(tcga_deconv) = colnames(tcga_deconv) %>% str_replace_all("(CBSX__sigmatrix_HNSCC_Fig2cd_)|(CBSX_sigmatrix_HNSCC_Fig2cd)", "CBSX.HNSCC")
colnames(tcga_deconv) = colnames(tcga_deconv) %>% str_replace_all("CBSX__LM22_" , "CBSX.LM22")

@@

Aran_all_facs$Sample_ID <- Aran_all_facs$Sample_ID %>% str_remove_all(., ".01A.*") %>% str_replace_all(., "-", ".")
Aran_all_facs <- Aran_all_facs[Aran_all_facs$CPE != "NaN",]

@@

Aran_all_facs <- Aran_all_facs %>% .[.$Sample_ID %in% Comm_samples, ] %>% .[order(.$Sample_ID), ]
HE_all_TCGA <- HE_all_TCGA %>% .[.$ParticipantBarcode %in% Comm_samples, ] %>% .[order(.$ParticipantBarcode),]
tcga_deconv <- tcga_deconv %>% .[.$Samples %in% Comm_samples, ]
tcga_deconv <- tcga_deconv

# REMOVE NON UNIQUE SAMPLES IN TCGA DECONV and attribute cancer type
tcga_deconv <- tcga_deconv %>%
  group_by(Samples) %>%
  summarise_all(mean)%>%
  as.data.frame()
tcga_deconv <- merge(tcga_deconv, Aran_all_facs[,1:2], by.x = "Samples", by.y = "Sample_ID")
rownames(tcga_deconv) = tcga_deconv$Samples
tcga_deconv <- tcga_deconv[,-1]

# for each method_signature make a DF with all 3 cell type in H&E
make_df <- function(met_sign, DF){
  data = DF %>% select(colnames(DF)[grepl(met_sign, colnames(DF))])
  lymph = data %>% 
    select(colnames(data)[grepl("CD4", colnames(data))], colnames(data)[grepl("CD8", colnames(data))], 
           colnames(data)[grepl("NK", colnames(data))]) %>% 
    replace(is.na(.), 0) %>%
    mutate(sum = rowSums(across(where(is.numeric))))
  macro = data %>% 
    select(colnames(data)[grepl("M0", colnames(data))], colnames(data)[grepl("M1", colnames(data))], 
           colnames(data)[grepl("M2", colnames(data))], colnames(data)[grepl("Mono", colnames(data))],
           colnames(data)[grepl("Macro", colnames(data))]) %>% 
    replace(is.na(.), 0) %>%
    mutate(sum = rowSums(across(where(is.numeric))))
  neutro = data %>% 
    select(colnames(data)[grepl("Neu", colnames(data))]) %>% 
    replace(is.na(.), 0) %>%
    mutate(sum = rowSums(across(where(is.numeric))))

  final_df = data.frame(Lymphocytes = lymph$sum, Neutrophils = neutro$sum,
                        Macrophages = macro$sum, row.names = rownames(data))
  final_df
  }

colcol = c("darkslategrey", "green", "coral")
make_plot <- function(data, method_data, method){
  ploty = list()
  for (i in 1:3) {
    ploty[[i]] <- local({
      i <- i
      x = ggplot(data, aes(x = data_HE[, i], y = method_data[, i])) +
        geom_point(colour = colcol[i]) +
        stat_smooth(method = lm, size = 0.2, color = "black", se = F, fullrange = T) +
        scale_x_continuous(limits = c(0, max(method_data[,i]))) +
        scale_y_continuous(limits = c(0, max(method_data[,i]))) +
        theme(axis.title.x= element_text(size = 12, face="bold"), 
            axis.title.y= element_text(size = 12, face="bold") )  + 
        ylab(method) + xlab("FACS") + ggtitle(colnames(method_data)[i]) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
        annotate("text", x = -Inf, y = Inf, hjust=-0.2, vjust=1, label = paste("italic(r)==", round(cor.test(method_data[, i], data_HE[, i])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
        annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=3, label = paste("RMSE =", round(sqrt(mean((method_data[, i] - data_HE[, i])^2, na.rm = T)), 2)), parse = F, size = 6) 
    })
  }
  grid.arrange(grobs = ploty, ncol = 3)
}

```

@@

data = tcga_deconv %>% filter(Cancer_type == "LUAD")
data_HE = HE_all_TCGA %>% filter(Study == "LUAD")
data_HE = data_HE[, 3:5]

QT = make_df("Quantiseq", data)
make_plot(data, QT, "QuantiSeq")

Epidish_BPRNACan = make_df("Epidish_BPRNACan", data)
make_plot(data, Epidish_BPRNACan, "Epidish_BPRNACan")

DeconRNASeq_BPRNACan = make_df("DeconRNASeq_BPRNACan", data)
make_plot(data, DeconRNASeq_BPRNACan, "DeconRNASeq_BPRNACan")

Epidish_BPRNACanProMet = make_df("Epidish_BPRNACanProMet", data)
make_plot(data, Epidish_BPRNACanProMet, "Epidish_BPRNACanProMet")
DeconRNASeq_BPRNACanProMet = make_df("DeconRNASeq_BPRNACanProMet", data)
make_plot(data, DeconRNASeq_BPRNACanProMet, "DeconRNASeq_BPRNACanProMet")

Epidish_BPRNACan3DProMet = make_df("Epidish_BPRNACan3DProMet", data)
make_plot(data, Epidish_BPRNACan3DProMet, "Epidish_BPRNACan3DProMet")
DeconRNASeq_BPRNACan3DProMet = make_df("DeconRNASeq_BPRNACan3DProMet", data)
make_plot(data, DeconRNASeq_BPRNACan3DProMet, "DeconRNASeq_BPRNACan3DProMet")

Epidish_CBSX.NSCLC = make_df("Epidish_CBSX.NSCLC", data)
make_plot(data, Epidish_CBSX.NSCLC, "Epidish_CBSX.NSCLC")
DeconRNASeq_CBSX.NSCLC = make_df("DeconRNASeq_CBSX.NSCLC", data)
make_plot(data, DeconRNASeq_CBSX.NSCLC, "DeconRNASeq_CBSX.NSCLC")

Epidish_CBSX_LM22 = make_df("Epidish_CBSX_LM22", data)
make_plot(data, Epidish_CBSX_LM22, "Epidish_CBSX_LM22")
DeconRNASeq_CBSX_LM22 = make_df("DeconRNASeq_CBSX_LM22", data)
make_plot(data, DeconRNASeq_CBSX_LM22, "DeconRNASeq_CBSX_LM22")

Epidish_CBSX.melanoma = make_df("Epidish_CBSX.melanoma", data)
make_plot(data, Epidish_CBSX.melanoma, "Epidish_CBSX.melanoma")
DeconRNASeq_CBSX.melanoma = make_df("DeconRNASeq_CBSX.melanoma", data)
make_plot(data, DeconRNASeq_CBSX.melanoma, "DeconRNASeq_CBSX.melanoma")

Epidish_CBSX.HNSCC = make_df("Epidish_CBSX.HNSCC", data)
make_plot(data, Epidish_CBSX.HNSCC, "Epidish_CBSX.HNSCC")
DeconRNASeq_CBSX.HNSCC = make_df("DeconRNASeq_CBSX.HNSCC", data)
make_plot(data, DeconRNASeq_CBSX.HNSCC, "DeconRNASeq_CBSX.HNSCC")

CBSX.LM22 = make_df("^CBSX.LM22", data)
make_plot(data, CBSX.LM22, "CBSX.LM22")
CBSX.NSCLC = make_df("^CBSX.NSCLC", data)
make_plot(data, CBSX.NSCLC, "CBSX.NSCLC")
CBSX.melanoma = make_df("^CBSX.melanoma", data)
make_plot(data, CBSX.melanoma, "CBSX.melanoma")
CBSX.HNSCC = make_df("^CBSX.HNSCC", data)
make_plot(data, CBSX.HNSCC, "CBSX.HNSCC")

```

@@

data = tcga_deconv %>% filter(Cancer_type == "LUSC")
data_HE = HE_all_TCGA %>% filter(Study == "LUSC")
data_HE = data_HE[, 3:5]

QT = make_df("Quantiseq", data)
make_plot(data, QT, "QuantiSeq")

Epidish_BPRNACan = make_df("Epidish_BPRNACan", data)
make_plot(data, Epidish_BPRNACan, "Epidish_BPRNACan")

DeconRNASeq_BPRNACan = make_df("DeconRNASeq_BPRNACan", data)
make_plot(data, DeconRNASeq_BPRNACan, "DeconRNASeq_BPRNACan")

Epidish_BPRNACanProMet = make_df("Epidish_BPRNACanProMet", data)
make_plot(data, Epidish_BPRNACanProMet, "Epidish_BPRNACanProMet")
DeconRNASeq_BPRNACanProMet = make_df("DeconRNASeq_BPRNACanProMet", data)
make_plot(data, DeconRNASeq_BPRNACanProMet, "DeconRNASeq_BPRNACanProMet")

Epidish_BPRNACan3DProMet = make_df("Epidish_BPRNACan3DProMet", data)
make_plot(data, Epidish_BPRNACan3DProMet, "Epidish_BPRNACan3DProMet")
DeconRNASeq_BPRNACan3DProMet = make_df("DeconRNASeq_BPRNACan3DProMet", data)
make_plot(data, DeconRNASeq_BPRNACan3DProMet, "DeconRNASeq_BPRNACan3DProMet")

Epidish_CBSX.NSCLC = make_df("Epidish_CBSX.NSCLC", data)
make_plot(data, Epidish_CBSX.NSCLC, "Epidish_CBSX.NSCLC")
DeconRNASeq_CBSX.NSCLC = make_df("DeconRNASeq_CBSX.NSCLC", data)
make_plot(data, DeconRNASeq_CBSX.NSCLC, "DeconRNASeq_CBSX.NSCLC")

Epidish_CBSX_LM22 = make_df("Epidish_CBSX_LM22", data)
make_plot(data, Epidish_CBSX_LM22, "Epidish_CBSX_LM22")
DeconRNASeq_CBSX_LM22 = make_df("DeconRNASeq_CBSX_LM22", data)
make_plot(data, DeconRNASeq_CBSX_LM22, "DeconRNASeq_CBSX_LM22")

Epidish_CBSX.melanoma = make_df("Epidish_CBSX.melanoma", data)
make_plot(data, Epidish_CBSX.melanoma, "Epidish_CBSX.melanoma")
DeconRNASeq_CBSX.melanoma = make_df("DeconRNASeq_CBSX.melanoma", data)
make_plot(data, DeconRNASeq_CBSX.melanoma, "DeconRNASeq_CBSX.melanoma")

Epidish_CBSX.HNSCC = make_df("Epidish_CBSX.HNSCC", data)
make_plot(data, Epidish_CBSX.HNSCC, "Epidish_CBSX.HNSCC")
DeconRNASeq_CBSX.HNSCC = make_df("DeconRNASeq_CBSX.HNSCC", data)
make_plot(data, DeconRNASeq_CBSX.HNSCC, "DeconRNASeq_CBSX.HNSCC")

CBSX.LM22 = make_df("^CBSX.LM22", data)
make_plot(data, CBSX.LM22, "CBSX.LM22")
CBSX.NSCLC = make_df("^CBSX.NSCLC", data)
make_plot(data, CBSX.NSCLC, "CBSX.NSCLC")
CBSX.melanoma = make_df("^CBSX.melanoma", data)
make_plot(data, CBSX.melanoma, "CBSX.melanoma")
CBSX.HNSCC = make_df("^CBSX.HNSCC", data)
make_plot(data, CBSX.HNSCC, "CBSX.HNSCC")

```

@@

data = tcga_deconv %>% filter(Cancer_type == "PRAD")
data_HE = HE_all_TCGA %>% filter(Study == "PRAD")
data_HE = data_HE[, 3:5]

QT = make_df("Quantiseq", data)
make_plot(data, QT, "QuantiSeq")

Epidish_BPRNACan = make_df("Epidish_BPRNACan", data)
make_plot(data, Epidish_BPRNACan, "Epidish_BPRNACan")

DeconRNASeq_BPRNACan = make_df("DeconRNASeq_BPRNACan", data)
make_plot(data, DeconRNASeq_BPRNACan, "DeconRNASeq_BPRNACan")

Epidish_BPRNACanProMet = make_df("Epidish_BPRNACanProMet", data)
make_plot(data, Epidish_BPRNACanProMet, "Epidish_BPRNACanProMet")
DeconRNASeq_BPRNACanProMet = make_df("DeconRNASeq_BPRNACanProMet", data)
make_plot(data, DeconRNASeq_BPRNACanProMet, "DeconRNASeq_BPRNACanProMet")

Epidish_BPRNACan3DProMet = make_df("Epidish_BPRNACan3DProMet", data)
make_plot(data, Epidish_BPRNACan3DProMet, "Epidish_BPRNACan3DProMet")
DeconRNASeq_BPRNACan3DProMet = make_df("DeconRNASeq_BPRNACan3DProMet", data)
make_plot(data, DeconRNASeq_BPRNACan3DProMet, "DeconRNASeq_BPRNACan3DProMet")

Epidish_CBSX.NSCLC = make_df("Epidish_CBSX.NSCLC", data)
make_plot(data, Epidish_CBSX.NSCLC, "Epidish_CBSX.NSCLC")
DeconRNASeq_CBSX.NSCLC = make_df("DeconRNASeq_CBSX.NSCLC", data)
make_plot(data, DeconRNASeq_CBSX.NSCLC, "DeconRNASeq_CBSX.NSCLC")

Epidish_CBSX_LM22 = make_df("Epidish_CBSX_LM22", data)
make_plot(data, Epidish_CBSX_LM22, "Epidish_CBSX_LM22")
DeconRNASeq_CBSX_LM22 = make_df("DeconRNASeq_CBSX_LM22", data)
make_plot(data, DeconRNASeq_CBSX_LM22, "DeconRNASeq_CBSX_LM22")

Epidish_CBSX.melanoma = make_df("Epidish_CBSX.melanoma", data)
make_plot(data, Epidish_CBSX.melanoma, "Epidish_CBSX.melanoma")
DeconRNASeq_CBSX.melanoma = make_df("DeconRNASeq_CBSX.melanoma", data)
make_plot(data, DeconRNASeq_CBSX.melanoma, "DeconRNASeq_CBSX.melanoma")

Epidish_CBSX.HNSCC = make_df("Epidish_CBSX.HNSCC", data)
make_plot(data, Epidish_CBSX.HNSCC, "Epidish_CBSX.HNSCC")
DeconRNASeq_CBSX.HNSCC = make_df("DeconRNASeq_CBSX.HNSCC", data)
make_plot(data, DeconRNASeq_CBSX.HNSCC, "DeconRNASeq_CBSX.HNSCC")

CBSX.LM22 = make_df("^CBSX.LM22", data)
make_plot(data, CBSX.LM22, "CBSX.LM22")
CBSX.NSCLC = make_df("^CBSX.NSCLC", data)
make_plot(data, CBSX.NSCLC, "CBSX.NSCLC")
CBSX.melanoma = make_df("^CBSX.melanoma", data)
make_plot(data, CBSX.melanoma, "CBSX.melanoma")
CBSX.HNSCC = make_df("^CBSX.HNSCC", data)
make_plot(data, CBSX.HNSCC, "CBSX.HNSCC")

```

@@

data = tcga_deconv %>% filter(Cancer_type == "BLCA")
data_HE = HE_all_TCGA %>% filter(Study == "BLCA")
data_HE = data_HE[, 3:5]

QT = make_df("Quantiseq", data)
make_plot(data, QT, "QuantiSeq")

Epidish_BPRNACan = make_df("Epidish_BPRNACan", data)
make_plot(data, Epidish_BPRNACan, "Epidish_BPRNACan")

DeconRNASeq_BPRNACan = make_df("DeconRNASeq_BPRNACan", data)
make_plot(data, DeconRNASeq_BPRNACan, "DeconRNASeq_BPRNACan")

Epidish_BPRNACanProMet = make_df("Epidish_BPRNACanProMet", data)
make_plot(data, Epidish_BPRNACanProMet, "Epidish_BPRNACanProMet")
DeconRNASeq_BPRNACanProMet = make_df("DeconRNASeq_BPRNACanProMet", data)
make_plot(data, DeconRNASeq_BPRNACanProMet, "DeconRNASeq_BPRNACanProMet")

Epidish_BPRNACan3DProMet = make_df("Epidish_BPRNACan3DProMet", data)
make_plot(data, Epidish_BPRNACan3DProMet, "Epidish_BPRNACan3DProMet")
DeconRNASeq_BPRNACan3DProMet = make_df("DeconRNASeq_BPRNACan3DProMet", data)
make_plot(data, DeconRNASeq_BPRNACan3DProMet, "DeconRNASeq_BPRNACan3DProMet")

Epidish_CBSX.NSCLC = make_df("Epidish_CBSX.NSCLC", data)
make_plot(data, Epidish_CBSX.NSCLC, "Epidish_CBSX.NSCLC")
DeconRNASeq_CBSX.NSCLC = make_df("DeconRNASeq_CBSX.NSCLC", data)
make_plot(data, DeconRNASeq_CBSX.NSCLC, "DeconRNASeq_CBSX.NSCLC")

Epidish_CBSX_LM22 = make_df("Epidish_CBSX_LM22", data)
make_plot(data, Epidish_CBSX_LM22, "Epidish_CBSX_LM22")
DeconRNASeq_CBSX_LM22 = make_df("DeconRNASeq_CBSX_LM22", data)
make_plot(data, DeconRNASeq_CBSX_LM22, "DeconRNASeq_CBSX_LM22")

Epidish_CBSX.melanoma = make_df("Epidish_CBSX.melanoma", data)
make_plot(data, Epidish_CBSX.melanoma, "Epidish_CBSX.melanoma")
DeconRNASeq_CBSX.melanoma = make_df("DeconRNASeq_CBSX.melanoma", data)
make_plot(data, DeconRNASeq_CBSX.melanoma, "DeconRNASeq_CBSX.melanoma")

Epidish_CBSX.HNSCC = make_df("Epidish_CBSX.HNSCC", data)
make_plot(data, Epidish_CBSX.HNSCC, "Epidish_CBSX.HNSCC")
DeconRNASeq_CBSX.HNSCC = make_df("DeconRNASeq_CBSX.HNSCC", data)
make_plot(data, DeconRNASeq_CBSX.HNSCC, "DeconRNASeq_CBSX.HNSCC")

CBSX.LM22 = make_df("^CBSX.LM22", data)
make_plot(data, CBSX.LM22, "CBSX.LM22")
CBSX.NSCLC = make_df("^CBSX.NSCLC", data)
make_plot(data, CBSX.NSCLC, "CBSX.NSCLC")
CBSX.melanoma = make_df("^CBSX.melanoma", data)
make_plot(data, CBSX.melanoma, "CBSX.melanoma")
CBSX.HNSCC = make_df("^CBSX.HNSCC", data)
make_plot(data, CBSX.HNSCC, "CBSX.HNSCC")

```

@@

data = tcga_deconv %>% filter(Cancer_type == "CESC")
data_HE = HE_all_TCGA %>% filter(Study == "CESC")
data_HE = data_HE[, 3:5]

QT = make_df("Quantiseq", data)
make_plot(data, QT, "QuantiSeq")

Epidish_BPRNACan = make_df("Epidish_BPRNACan", data)
make_plot(data, Epidish_BPRNACan, "Epidish_BPRNACan")

DeconRNASeq_BPRNACan = make_df("DeconRNASeq_BPRNACan", data)
make_plot(data, DeconRNASeq_BPRNACan, "DeconRNASeq_BPRNACan")

Epidish_BPRNACanProMet = make_df("Epidish_BPRNACanProMet", data)
make_plot(data, Epidish_BPRNACanProMet, "Epidish_BPRNACanProMet")
DeconRNASeq_BPRNACanProMet = make_df("DeconRNASeq_BPRNACanProMet", data)
make_plot(data, DeconRNASeq_BPRNACanProMet, "DeconRNASeq_BPRNACanProMet")

Epidish_BPRNACan3DProMet = make_df("Epidish_BPRNACan3DProMet", data)
make_plot(data, Epidish_BPRNACan3DProMet, "Epidish_BPRNACan3DProMet")
DeconRNASeq_BPRNACan3DProMet = make_df("DeconRNASeq_BPRNACan3DProMet", data)
make_plot(data, DeconRNASeq_BPRNACan3DProMet, "DeconRNASeq_BPRNACan3DProMet")

Epidish_CBSX.NSCLC = make_df("Epidish_CBSX.NSCLC", data)
make_plot(data, Epidish_CBSX.NSCLC, "Epidish_CBSX.NSCLC")
DeconRNASeq_CBSX.NSCLC = make_df("DeconRNASeq_CBSX.NSCLC", data)
make_plot(data, DeconRNASeq_CBSX.NSCLC, "DeconRNASeq_CBSX.NSCLC")

Epidish_CBSX_LM22 = make_df("Epidish_CBSX_LM22", data)
make_plot(data, Epidish_CBSX_LM22, "Epidish_CBSX_LM22")
DeconRNASeq_CBSX_LM22 = make_df("DeconRNASeq_CBSX_LM22", data)
make_plot(data, DeconRNASeq_CBSX_LM22, "DeconRNASeq_CBSX_LM22")

Epidish_CBSX.melanoma = make_df("Epidish_CBSX.melanoma", data)
make_plot(data, Epidish_CBSX.melanoma, "Epidish_CBSX.melanoma")
DeconRNASeq_CBSX.melanoma = make_df("DeconRNASeq_CBSX.melanoma", data)
make_plot(data, DeconRNASeq_CBSX.melanoma, "DeconRNASeq_CBSX.melanoma")

Epidish_CBSX.HNSCC = make_df("Epidish_CBSX.HNSCC", data)
make_plot(data, Epidish_CBSX.HNSCC, "Epidish_CBSX.HNSCC")
DeconRNASeq_CBSX.HNSCC = make_df("DeconRNASeq_CBSX.HNSCC", data)
make_plot(data, DeconRNASeq_CBSX.HNSCC, "DeconRNASeq_CBSX.HNSCC")

CBSX.LM22 = make_df("^CBSX.LM22", data)
make_plot(data, CBSX.LM22, "CBSX.LM22")
CBSX.NSCLC = make_df("^CBSX.NSCLC", data)
make_plot(data, CBSX.NSCLC, "CBSX.NSCLC")
CBSX.melanoma = make_df("^CBSX.melanoma", data)
make_plot(data, CBSX.melanoma, "CBSX.melanoma")
CBSX.HNSCC = make_df("^CBSX.HNSCC", data)
make_plot(data, CBSX.HNSCC, "CBSX.HNSCC")

```

@@

data = tcga_deconv %>% filter(Cancer_type == "BRCA")
data_HE = HE_all_TCGA %>% filter(Study == "BRCA")
data_HE = data_HE[, 3:5]

QT = make_df("Quantiseq", data)
make_plot(data, QT, "QuantiSeq")

Epidish_BPRNACan = make_df("Epidish_BPRNACan", data)
make_plot(data, Epidish_BPRNACan, "Epidish_BPRNACan")

DeconRNASeq_BPRNACan = make_df("DeconRNASeq_BPRNACan", data)
make_plot(data, DeconRNASeq_BPRNACan, "DeconRNASeq_BPRNACan")

Epidish_BPRNACanProMet = make_df("Epidish_BPRNACanProMet", data)
make_plot(data, Epidish_BPRNACanProMet, "Epidish_BPRNACanProMet")
DeconRNASeq_BPRNACanProMet = make_df("DeconRNASeq_BPRNACanProMet", data)
make_plot(data, DeconRNASeq_BPRNACanProMet, "DeconRNASeq_BPRNACanProMet")

Epidish_BPRNACan3DProMet = make_df("Epidish_BPRNACan3DProMet", data)
make_plot(data, Epidish_BPRNACan3DProMet, "Epidish_BPRNACan3DProMet")
DeconRNASeq_BPRNACan3DProMet = make_df("DeconRNASeq_BPRNACan3DProMet", data)
make_plot(data, DeconRNASeq_BPRNACan3DProMet, "DeconRNASeq_BPRNACan3DProMet")

Epidish_CBSX.NSCLC = make_df("Epidish_CBSX.NSCLC", data)
make_plot(data, Epidish_CBSX.NSCLC, "Epidish_CBSX.NSCLC")
DeconRNASeq_CBSX.NSCLC = make_df("DeconRNASeq_CBSX.NSCLC", data)
make_plot(data, DeconRNASeq_CBSX.NSCLC, "DeconRNASeq_CBSX.NSCLC")

Epidish_CBSX_LM22 = make_df("Epidish_CBSX_LM22", data)
make_plot(data, Epidish_CBSX_LM22, "Epidish_CBSX_LM22")
DeconRNASeq_CBSX_LM22 = make_df("DeconRNASeq_CBSX_LM22", data)
make_plot(data, DeconRNASeq_CBSX_LM22, "DeconRNASeq_CBSX_LM22")

Epidish_CBSX.melanoma = make_df("Epidish_CBSX.melanoma", data)
make_plot(data, Epidish_CBSX.melanoma, "Epidish_CBSX.melanoma")
DeconRNASeq_CBSX.melanoma = make_df("DeconRNASeq_CBSX.melanoma", data)
make_plot(data, DeconRNASeq_CBSX.melanoma, "DeconRNASeq_CBSX.melanoma")

Epidish_CBSX.HNSCC = make_df("Epidish_CBSX.HNSCC", data)
make_plot(data, Epidish_CBSX.HNSCC, "Epidish_CBSX.HNSCC")
DeconRNASeq_CBSX.HNSCC = make_df("DeconRNASeq_CBSX.HNSCC", data)
make_plot(data, DeconRNASeq_CBSX.HNSCC, "DeconRNASeq_CBSX.HNSCC")

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(reshape2)
  library(ggpubr)
  library(tidyverse)
  library(Hmisc)
  library(ggridges)
  library(gridExtra)
  library(grid)
})
knitr::opts_chunk$set(fig.width = 15, fig.height = 15)
```

@@

tcga_deconv <- read.table("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/deconvolutions_pipeline/all_deconvolutions/all_deconvolutions_all_TCGA.txt", header = T)
Aran_all_facs <- read.csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Aran_facs.csv", sep = "\t")
HE_all_TCGA <- read.csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv", row.names = 1)

colnames(tcga_deconv) <- colnames(tcga_deconv) %>% str_replace_all("(CBSX__Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix_)|(CBSX_Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix)", "CBSX.NSCLC")
colnames(tcga_deconv) <- colnames(tcga_deconv) %>% str_replace_all("(CBSX__scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b_)|(CBSX_scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b)", "CBSX.melanoma")
colnames(tcga_deconv) <- colnames(tcga_deconv) %>% str_replace_all("(CBSX__sigmatrix_HNSCC_Fig2cd_)|(CBSX_sigmatrix_HNSCC_Fig2cd)", "CBSX.HNSCC")
colnames(tcga_deconv) <- colnames(tcga_deconv) %>% str_replace_all("CBSX__LM22_", "CBSX.LM22")

@@

Aran_all_facs$Sample_ID <- Aran_all_facs$Sample_ID %>%
  str_remove_all(., ".01A.*") %>%
  str_replace_all(., "-", ".")
Aran_all_facs <- Aran_all_facs[Aran_all_facs$CPE != "NaN", ]

@@

Aran_all_facs <- Aran_all_facs %>%
  .[.$Sample_ID %in% Comm_samples, ] %>%
  .[order(.$Sample_ID), ]
HE_all_TCGA <- HE_all_TCGA %>%
  .[.$ParticipantBarcode %in% Comm_samples, ] %>%
  .[order(.$ParticipantBarcode), ]
tcga_deconv <- tcga_deconv %>% .[.$Samples %in% Comm_samples, ]
tcga_deconv <- tcga_deconv

# REMOVE NON UNIQUE SAMPLES IN TCGA DECONV and attribute cancer type
tcga_deconv <- tcga_deconv %>%
  group_by(Samples) %>%
  summarise_all(mean) %>%
  as.data.frame()
tcga_deconv <- merge(tcga_deconv, Aran_all_facs[, 1:2], by.x = "Samples", by.y = "Sample_ID")
rownames(tcga_deconv) <- tcga_deconv$Samples
tcga_deconv <- tcga_deconv[, -1]

# for each method_signature make a DF with all 3 cell type in H&E
make_df <- function(met_sign, DF) {
  data <- DF %>% select(colnames(DF)[grepl(met_sign, colnames(DF))])
  lymph <- data %>%
    select(
      colnames(data)[grepl("CD4", colnames(data))], colnames(data)[grepl("CD8", colnames(data))],
      colnames(data)[grepl("NK", colnames(data))]
    ) %>%
    replace(is.na(.), 0) %>%
    mutate(sum = rowSums(across(where(is.numeric))))
  macro <- data %>%
    select(
      colnames(data)[grepl("M0", colnames(data))], colnames(data)[grepl("M1", colnames(data))],
      colnames(data)[grepl("M2", colnames(data))], colnames(data)[grepl("Mono", colnames(data))],
      colnames(data)[grepl("Macro", colnames(data))]
    ) %>%
    replace(is.na(.), 0) %>%
    mutate(sum = rowSums(across(where(is.numeric))))
  neutro <- data %>%
    select(colnames(data)[grepl("Neu", colnames(data))]) %>%
    replace(is.na(.), 0) %>%
    mutate(sum = rowSums(across(where(is.numeric))))

  final_df <- data.frame(
    Lymphocytes = lymph$sum, Neutrophils = neutro$sum,
    Macrophages = macro$sum, row.names = rownames(data)
  )
  final_df
}

colcol <- c("darkslategrey", "green", "coral")
make_plot <- function(data, data_HE, method, method_data, cancer_type) {
  ploty <- list()
  for (i in 1:3) {
    ploty[[i]] <- local({
      i <- i
      ggplot(data, aes(x = data_HE[, i], y = method_data[, i])) +
        geom_point(colour = colcol[i]) +
        stat_smooth(method = lm, formula = y ~ x, size = 0.2, color = "black", se = F, fullrange = T) +
        scale_x_continuous(limits = c(0, max(method_data[, i]))) +
        scale_y_continuous(limits = c(0, max(method_data[, i]))) +
        theme(
          axis.title.x = element_text(size = 12, face = "bold"),
          axis.title.y = element_text(size = 12, face = "bold")
        ) +
        ylab(method) +
        xlab("FACS") +
        ggtitle(colnames(method_data)[i]) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
        annotate("text", x = -Inf, y = Inf, hjust = -0.2, vjust = 1, label = paste("italic(r)==", round(cor.test(method_data[, i], data_HE[, i])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
        annotate("text", x = -Inf, y = Inf, hjust = -1, vjust = 3, label = paste("RMSE =", round(sqrt(mean((method_data[, i] - data_HE[, i])^2, na.rm = T)), 2)), parse = F, size = 6)
    })
  }
  grid.arrange(grobs = ploty, ncol = 3)
}

make_plots <- function(tcga_deconv, signatures_list, HE_all_TCGA, cancer_type) {
  cancer_type_data <- tcga_deconv %>% filter(Cancer_type == cancer_type)

  cancer_type_data_HE <- HE_all_TCGA %>% filter(Study == cancer_type) %>% select(3:5)

  for (signature in signatures_list) {
    signature_data <- make_df(signature, cancer_type_data)
    curated_signature <- str_remove(signature, fixed("^"))
    make_plot(cancer_type_data, cancer_type_data_HE, curated_signature, signature_data, cancer_type)
  }
}
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
signatures_list <- c("Quantiseq", "Epidish_BPRNACan", "DeconRNASeq_BPRNACan", "Epidish_BPRNACanProMet", "DeconRNASeq_BPRNACanProMet", "Epidish_BPRNACan3DProMet", "DeconRNASeq_BPRNACan3DProMet", "Epidish_CBSX.NSCLC", "DeconRNASeq_CBSX.NSCLC", "Epidish_CBSX_LM22", "DeconRNASeq_CBSX_LM22", "Epidish_CBSX.melanoma", "DeconRNASeq_CBSX.melanoma", "Epidish_CBSX.HNSCC", "DeconRNASeq_CBSX.HNSCC", "^CBSX.LM22", "^CBSX.NSCLC", "^CBSX.melanoma", "^CBSX.HNSCC")
```

@@

make_plots(tcga_deconv, signatures_list, HE_all_TCGA, "LUAD")
```

@@

make_plots(tcga_deconv, signatures_list, HE_all_TCGA, "LUSC")
```

@@

make_plots(tcga_deconv, signatures_list, HE_all_TCGA, "PRAD")
```

@@

make_plots(tcga_deconv, signatures_list, HE_all_TCGA, "BLCA")
```

@@

make_plots(tcga_deconv, signatures_list, HE_all_TCGA, "CESC")
```

@@

make_plots(tcga_deconv, signatures_list, HE_all_TCGA, "BRCA")
```
